---
title: "apply_functions"
output:
  pdf_document: default
  html_document: default
date: "2023-04-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
```

```{r}
?lapply
?mapply
```
The `apply` functions allow us to apply a function to a vector or list of values iteratively.
This helps minimize errors in code and makes the analyses more efficient.
 With lapply() and sapply functions, we can only provide one argument to iterate on 
 
 With mapply(), we can provide multiple arguments to iterate. It probably also returns a vector or simpflied data structure as result.

`sapply()` function simpflies the output to a vector (or the simplest data structure possible), while `lapply()` returns an output in the form of a list. 

```{r}
get_mass_from_length_theropoda <- function(length){
  mass <- 0.73 * length ^ 3.63
  return(mass)
}
```

```{r}
theropoda_lengths <- c(17.8013631070471, 20.3764452071665, 14.0743486294308, 25.65782386974, 26.0952008049675, 20.3111541103134, 17.5663244372533, 11.2563431277577, 20.081903202614, 18.6071626441984, 18.0991894513166, 23.0659685685892, 20.5798853467837, 25.6179254233558, 24.3714331573996, 26.2847248252537, 25.4753783544473, 20.4642089867304, 16.0738256364701, 20.3494171706583, 19.854399305869, 17.7889814608919, 14.8016421998303, 19.6840911485379, 19.4685885050906, 24.4807784966691, 13.3359960054899, 21.5065994598917, 18.4640304608411, 19.5861532398676, 27.084751999756, 18.9609366301798, 22.4829168046521, 11.7325716149514, 18.3758846100456, 15.537504851634, 13.4848751773738, 7.68561192214935, 25.5963348603783, 16.588285389794)
```

```{r}
get_mass_from_length_theropoda(length = theropoda_lengths)
```

```{r}
theropoda_masses <- get_mass_from_length_theropoda(length = theropoda_lengths)

mylist <- (theropoda_masses)
second_list <- c(mylist, list(c("Luna", "Avi","Anita")))
second_list[[1]]

data.frame(theropoda_masses, c("Anita", "Avi", "Luna", "Maria"))
```



```{r}
mass_from_length <- function(length, a,b){
  mass <- 0.73 * length ^ 3.63
  return(mass)
}
new_masses <-mass_from_length(length = theropoda_lengths)
#rm(new_lengths) # The rm function allows to remove objects from the R environment.

theropoda_masses == new_masses
all(theropoda_masses == new_masses) # tests that all values in a logica vector are equal to TRUE
all.equal(theropoda_masses, new_masses)
```

```{r}
a_values <- c(0.759, 0.751, 0.74, 0.746, 0.759, 0.751, 0.749, 0.751, 0.738, 0.768, 0.736, 0.749, 0.746, 0.744, 0.749, 0.751, 0.744, 0.754, 0.774, 0.751, 0.763, 0.749, 0.741, 0.754, 0.746, 0.755, 0.764, 0.758, 0.76, 0.748, 0.745, 0.756, 0.739, 0.733, 0.757, 0.747, 0.741, 0.752, 0.752, 0.748)
```

```{r}
b_values <- c(3.627, 3.633, 3.626, 3.633, 3.627, 3.629, 3.632, 3.628, 3.633, 3.627, 3.621, 3.63, 3.631, 3.632, 3.628, 3.626, 3.639, 3.626, 3.635, 3.629, 3.642, 3.632, 3.633, 3.629, 3.62, 3.619, 3.638, 3.627, 3.621, 3.628, 3.628, 3.635, 3.624, 3.621, 3.621, 3.632, 3.627, 3.624, 3.634, 3.621)
```

```{r}
mass_from_length <- function(length = theropoda_lengths, a = a_values, b=  b_values){
  mass <- a * length ^ b
  return(mass)
}
```

`<<-` the scope operator or double arrow, allows creating and modifying objects in parent variables

```{r}
dino_data <- data.frame(theropoda_lengths, a_values, b_values) %>% mutate(massses = get_mass_from_length_theropoda(theropoda_lengths)) 
print(dino_data)
```

```{r error=TRUE}

theropoda_lengths < 20

mass_from_length_max <- function(length) {
  if(length < 20) {
    mass <- 0.73 * length ^ 3.63
    return(mass)
    } else {
      return(NA)
    }
    }

    
    mass_from_length_max <- function(length) {
  if(length < 20) {
    mass <- 0.73 * length ^ 3.63
    mass <-NA
    }
    }
    
    
mass_from_length_max <- function(length) {
  if(length < 20) {
    mass <- 0.73 * length ^ 3.63
    return(mass)
  }
}

```

```{r error=TRUE}
get_mass_from_length_theropoda(length = theropoda_lengths)
```

```{r}
sapply(theropoda_lengths, mass_from_length_max)
```


#Homework
```{r}
dinosaur_lengths <- read.csv(file = "../data raw/dinosaur_lengths.csv")
head(dinosaur_lengths)
```

```{r}
get_mass_from_length_by_name <- function(length, dinosaur_name) {
if (dinosaur_name == "Stegosauria") {
a <- 10.95
b <- 2.64
} else if (dinosaur_name == "Theropoda") {
a <- 0.73
b <- 3.63
} else if (dinosaur_name == "Sauropoda") {
a <- 214.44
b <- 1.46 
} else {
a = NA
b = NA
}
mass <- a * length^b
return(mass)
}
```

```{r}
mapply(get_mass_from_length_by_name, length = dinosaur_lengths$lengths, dinosaur_name = dinosaur_lengths$species)
```

```{r}
dinosaur_lengths %>% 
rowwise %>% 
mutate(masses = get_mass_from_length_by_name(lengths, species))-> dino_masses
  head(dino_masses) 
```

```{r}
library("ggplot2")
ggplot(data = dino_masses, mapping = aes(x = masses)) +
geom_histogram() +
facet_wrap(~species)

head(dinosaur_lengths)


dino_masses %>%
  filter(!is.na(masses)) %>% 
  ggplot() +
  geom_histogram(mapping = aes(x = masses, color = species)) +
  facet_wrap(~species)
```


