---
title: "functions.rmd"
output:
  pdf_document: default
  html_document: default
date: "2023-03-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

```{r}
get_mass_from_length_theropoda <- function(length){
  mass <- 0.73 * length ^ 3.63
  return(mass)
}
```


##Writing function - Exercise 1
```{r}
pounds <- 3
pounds = 3
yellow <- "Avi"
red <- 3.75
convert_pounds_to_grams <- function(pounds) {
     grams = 453.6 * pounds
     return(grams)
}

convert_pounds_to_grams <- function(pounds) {
     grams = 3.75 * pounds
     return(grams)
}

convert_pounds_to_grams(pounds = 1300)
length(x= yellow)


#a)   convert_pounds_to_grams - This is the object
#b) { grams = 453.6 * pounds  return(________) } - #It is the last output of a function.
#c) #function(________) - It is the argument
```



```{r}
get_mass_from_length_theropoda(16)
```


```{r}
get_mass_from_length <- function(length, a, b) {
  mass <- a * length ^ b
  return(mass)
}
```


```{r}
get_mass_from_length(length = 26, a = 214.44, b = 1.46)
```

## Why do we use functions in programming?

- We already use functions a lot. Basically everything we run in R is a function. 
- Reduce the amount of code we write.
- No need to repeat code, so less chance of repeating errors or introducing new errors by mistyping code.
- Helps with reproducibility: no need to repeat code, allows others to run code in a straightforward way, with accuracy.
- Makes it easier to write modular code that you can reuse for other projects. 
- Makes it easier to run code and remember it. 
- It allows to organize analysis better, by putting code intended to run something together in the same function. 

## Combining Functions Exercise
```{r}
kg = 5
convert_kg_to_pounds <- function(kg= 5) {
  pounds = 2.205 * kg
  return(pounds)
}

convert_kg_to_pounds(kg = 5)
convert_kg_to_pounds(50)

get_mass_from_length(length = 12, a = 10.95, b = 2.64) %>% 
  convert_kg_to_pounds()


length_in_pounds <- function(length, a, b){
  get_mass_from_length(length = length, a = 10.95, b = 2.64) %>% 
  convert_kg_to_pounds()
}
get_mass_from_length(length = 12, a = 10.95, b = 2.64)

```
```{r}
length_in_pounds(length = 35, a = 13, b = 1.3)
```

