---
title: "if.else"
output:
  pdf_document: default
  html_document: default
date: "2023-04-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Logical and conditional statements

These are the pieces of code that return the `TRUE` or `FALSE` values, that is, a logical value.

The common operators of logical statements are:
- equality `==`
- inequality `!=`
- greater than `>`
- less than `<`
- greater or equal then `>=`
- less or equal than `=`

The conditional statements allow to test several logical conditions at a time 
The condition operators (or symbols) are 
- AND `&` (inside dplyr function we can represent AND using a `,`)
- OR `|` 

We also have logical functions that test if something is `TRUE` or `FALSE`, for example:
- is.na()` is a function that tests if a value os an `NA`
-This function i part of a whole family of functions, they all start with `is.`:
is.vector()
is.data.frame()
- is.factor()

For next class: how to get all functions from a family (method).

-which(): takes logical vectors, it will give you a numerical index(position) of all values that are TRUE

```{r}
which(letters == "r")
letters[18]
```
```{r}
w <- 10.2
 x <- 1.3
 y <- 2.8
 z <- 17.5
 colors <- c("red", "blue", "green")
 masses <- c(45.2, 36.1, 27.8, 81.6, 42.4)
 dna1 <- "attattaggaccaca"
 dna2 <- "attattaggaacaca"

```


```{r}
w > 10
colors = "green"
x > y
masses > 40
(2 * x + 0.2) == y
dna1 = dna2
dna1 != dna2
w > x
x * w < 13.5
x * w > 13.2 
#13.2 < x * w < 13.5 This is how we do it in paper
#but in R we have to compare things in pairs:
#for this we use the conditional statements
x * w < 13.5 & x * w > 13.2
masses < 30 & 50 > 30 & 50
```
## How to make simple choices with `if`()`

```
The general structure of an if statement:
if (condition is TRUE) {
RUN all lines 
of code in 
this block 
of code
}
```

#If the condition is not TRUE, then nothing happens.

```{r}
age_class = "sapling"
if((age_class == "sapling")) {
y <- 10
}
y
```
## Case when we have two options: if-else structure

The general form of this structure:
```
if(condition){
} else{
code that runs if condition is NOT met
}
```
```{r}
age_class = "seedling"
if((age_class == "sapling")) {
} else
y <- 5
```

```{r}
if((age_class == "sapling")) {
y <- 10
}
y
 if((age_class == "sapling")) {
y <- 10
 }else{
 y <-5
}
y
```
```{r}
if (age_class == "seedling") {
y <-5
} else {
y <-10
}
```

### Handle more than 2 choices
In this case we are using the elseif structure:

```
if(condition1){
first block code that is executes if condition 1 is met
} else if (condition2) {
second block code that executes if condition2 is met 
} else if (condition3) {
more code
} else {
this will cover all the conditions that are not specified before
}

You do not have to end up with and else block. 
Else if are more intentionl with the conditions.
A simple `else` will run 
}
}
}
```
Handling more than 3 choices
```{r}
if((age_class == "sapling")) {
 y <- 10
} else if(age_class == "seedling") {
 y <-5
} else if((age_class == "adult")) {
 y <- 20
}
```

#Values of y by age class
```{r}

```