---
title: "joining-vectors"
output:
  pdf_document: default
  html_document: default
date: "2023-03-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```
library(dyplr)

## What are vectors?

- they are unidimensional matrices
- they can only hold one type of data, either numeric (integer or double), character, or logical(complex numbers)

##What are data frames

- It is a two dimensional matrix, rows, and columns 
- it can hold any type of data
- it can only hold different of data in a certain way"
- only columns can have differnt data types
- within a column, all rows have the same data type

- A data frame can also be defined as a colection of vectors (they can be of different or the same type) all of the same length!


```{r}
surveys <- read.csv(file = "../data raw/surveys.csv")
species <- read.csv (file = "../data raw/species.csv")
str(surveys)
```

## Creating vectors - review
```{r}
c("luna", "Avi", "Anita", "James", "Charles", "Damian", "Davinder") -> our_names
1:7 # the colon operator creates a vector of numbers
1:7 -> my_numbers
-100:200
```


#In how steps does the colon operator increase? It increases in a unit of 1.

What do we do if I want to create a numeric sequence that increases in steps different than 1?
```{r}
seq
seq(-100,200, by = 0.1) -> my_numbers
str(my_numbers)
```

```{r}
letters
LETTERS
```

## Creating data frames from vectors

The most general way to do this is with teh function `data.frame()` :
```{r error=TRUE}
data.frame(names = our_names, numbers = my_numbers)
```
Remember: Vectors must have the same length (or be a multiple) to be part of a data frame!

```{r}
data.frame(names = our_names, numbers =1:7)
```
R will only recycle the values only if they are multipe of the vector:

```{r}
data.frame(names = our_names, numbers = 1)
```
To recycle the values of a numeric vector of length 2, we have to repeat the vector of names two times, so it is a multiple of 2.
```{r}
data.frame(names = rep(our_names, 2), numbers = c(2, 5.5))
```
You have data on the length, width, and height of 10 individuals of the yew Taxus baccata stored in the following vectors:
Make a data frame that contains these three vectors as columns along with a "genus" column containing the genus name Taxus on all rows and a "species" column containing the species epithet baccata on all rows.

```{r}
length <- c(2.2, 2.1, 2.7, 3.0, 3.1, 2.5, 1.9, 1.1, 3.5, 2.9)
width <- c(1.3, 2.2, 1.5, 4.5, 3.1, NA, 1.8, 0.5, 2.0, 2.7)
height <- c(9.6, 7.6, 2.2, 1.5, 4.0, 3.0, 4.5, 2.3, 7.5, 3.2)
str(species)
data.frame(names = length, width, height, genus = "Taxus", species = "baccata")
```
## Extracting/accessing values from vectors and data frames
```{r}
surveys <- read.csv(file = "../data raw/surveys.csv")
species <- read.csv (file = "../data raw/species.csv")

str(surveys)
```

One common way to extract or access vectors from column in a data frame is the dollar sign symbol `$`

```{r}
surveys$record_id -> record_id
```

Another way is with the square brackets `[]`
```{r}
surveys[1:10, "hindfoot_length"]
```

If I want all the values from the rows of column hindfoot length:
```{r}
surveys[, "hindfoot_length"] -> hindfoot_length
```

Another way is to use double square brackets
```{r}
surveys[["record_id"]] %>% head()
```

```{r}
surveys$weight -> surveys_weight
surveys [,"month"] -> surveys_month
surveys$hindfoot_length -> hindfoot_length
str(hindfoot_length)

na.omit(hindfoot_length) %>% 
  mean
```

