---
title: "Tree from Open Tree"
output: html_document
date: "2023-04-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ape) # Analysis, Phylogenetics , and Evolution R package
#install.packages("rotl")
library(rotl) # an R interface for the Open Tree of Life

```

Insects- 1 million named species
Mammals - 5k named species
Flowering plants: 200k named species

### Read a newick tree from the Open Tree of Life'

The following function can only read trees in newick format. Make sure your file is a newick file. 
If you got your tree from the Phylesystem 

```{r}
tree <- read.tree(file = "../data raw/Weigmann_2009.tre")
```

Now plot the tree:
```{r, fig.height= 10, fig.width =10 }
plot(tree, cex = 0.6)
mtext(text = "Weigmann 2009 OpenTree tree")
```

#We can also use a url of a tree to read a tree into R. We will provide the URL in the argument `file =`
In this case, we do nor have to keep a copy of the file, we can just download it directly
The url is “http://ape-package.ird.fr/APER/APER2/primfive.tre”:

```{r}
?read.tree
small_tree <- read.tree(file = "http://ape-package.ird.fr/APER/APER2/primfive.tre")
```

The R structure of a phylogenetic tree

### Getting a tree for your species from the Open Tree of Life


The Open Tree of Like has tools that allow to match the names of a group or species to there unfied taxonomy so that we can look for information on their databases. 

If we are doing this in R, we are using the `tnrs_match_names()` from the package `rotl`"; tnrs stands for taxonomic name resolution service:

```{r}
tnrs <- tnrs_match_names(names = "amphibians")
class(tnrs)
```

What is the difference we have in `search_string` and `unique_name`

One is the search string and the other is the scientific taxonomic name from the Open Tree of Life unified Taxonomy (OTT).


We can use the unique number identifier from OTT to get information for that taxon. 

To get a tree we can use the function `tol_subtree()`

```{r}
tol_subtree(ott_id = tnrs$ott_id)
```

Homework: Try to get OTT id for my organism
